---
title: "Urchins"
author: "Jacob Weverka"
date: "11/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(janitor)

```


Read in urchin data
```{r message=FALSE}

urchins = read_csv(here("data/urchins.csv")) %>%
  clean_names()


```


###Size comparison of urchins

```{r}
ggplot(urchins, aes(x = common_name, y = size)) +
  geom_boxplot() +
  xlab("Common Name") +
  ylab("Size")

ggplot(urchins, aes(x = size)) +
  geom_histogram(binwidth = 0.5) +
  facet_wrap(~common_name) +
  xlab("Size") +
  ylab("Count")
```
```{r message=FALSE}

urchin_sum = urchins %>% 
  group_by(common_name) %>% 
  summarize(mean = mean(size), median = median(size), sd = sd(size), n = n())

kableExtra::kable(urchin_sum)

```

Red urchins appear to have a greater mean size than purple urchins. Red urchins also appear to more variable in size. All the largest individuals sampled were red urchins.


### Purple Urchins

Here, we're assigning proper reef names to each site, fixing dates, and summarizing the count of purple urchins for each reef in each year.

```{r}

purple = urchins %>%
  filter(common_name == "Purple Urchin") %>% 
  mutate(date = lubridate::mdy(date),
         site_full = case_when(
           site == "NAPL" ~ "Naples Reef",
           site == "AQUE" ~ "Arroyo Quemado Reef",
           site == "CARP" ~ "Carpinteria Reef",
           site == "MOHK" ~ "Mohawk Reef",
           site == "IVEE" ~ "Isla Vista Reef"
         )) %>% 
  mutate(year = as.integer(lubridate::year(date)))


purple_count = purple %>%
  group_by(site_full, year) %>%
  count()


```



Plotting the change in purple urchin abundance on each reef through time.

```{r}
ggplot(purple_count, aes(x = year, y = n, color = site_full)) +
  geom_line(size = 1.5) +
  scale_x_continuous(breaks=seq(2007, 2020, 1))
```
































